Assessing Gender Bias in Machine Translation

### 2017-12-14 paper/main.tex
Updated scripts, paper, tables, languages, jobs and added julia script to generate plots

\usepackage{multirow}
\usepackage{hyperref}
\usepackage{float}
\usepackage{array}
\usepackage{dirtytalk}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|m{3cm}|c|}
	\hline
	Language Family & Language & Pronominal Gender System & Supported \\ \hline
	\multirow{2}{*}{Austronesian} & Malay & No & Yes \\
	& Tagalog & No & \textbf{No} \\ \hline
	\multirow{3}{*}{Finno-Ugric} & Estonian & No & Yes \\
	& Finnish & No & Yes\\
	& Hungarian & No & Yes \\ \hline
	\multirow{4}{*}{Indo-European} & Armenian & No & Yes \\
	& Bengali & No & Yes \\
	& English & \textbf{Yes} & Yes \\
	& Persian & \textbf{Yes} & Yes \\ \hline
	\multirow{3}{*}{Indo-Aryan} & Maithili & No & \textbf{No} \\
	& Nepali & No & Yes \\
	& Oriya & No & \textbf{No} \\ \hline
	\multirow{10}{*}{} & Japanese & No & Yes \\
	& Korean & \textbf{Optional} & Yes \\
	& Turkish & No & Yes \\
	& Yoruba & No & Yes \\
	& Basque & No & Yes \\
	& Swahili & No & Yes \\
	& Chinese & \textbf{Optional} & Yes \\
	& Cantonese & No & \textbf{No} \\
	& Pipil & No & \textbf{No} \\
	& Quechuan & No & \textbf{No} \\ \hline
	\multirow{4}{*}{Constructed} & Esperanto & \textbf{Yes} & Yes \\
	& Ido & \textbf{Optional} & \textbf{No} \\
	& Lingua Franca Nova & No & \textbf{No} \\
	& Interlingua & No & \textbf{No} \\ \hline
	\end{tabular}
	\caption{Selected gender neutral languages obtained from the Wikipedia article \url{en.wikipedia.org/wiki/Gender_neutrality_in_genderless_languages}. Languages are grouped according to language families and classified according to whether they exhibit pronominal gender system. For the purposes of this work, we have decided to work only with languages lacking such a system, and as such Persian and Esperanto have been discarded. Because Tagalog, Maithili, Oriya, Cantonese, Pipil, Quechuan, Lingua Franca Nova and Interlingua lack support from Google Translate, these languages have also been ommitted from this work.}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
	\hline
	Category 		& \# Occupations 	\\ \hline
	Artistic 		& $102$ 			\\ \hline
	Computer 		& $19$ 				\\ \hline
	Corporate 		& $50$ 				\\ \hline
	Dance 			& $9$ 				\\ \hline
	Film/Television & $26$ 				\\ \hline
	Healthcare 		& $88$ 				\\ \hline
	Industrial 		& $26$ 				\\ \hline
	Science 		& $50$ 				\\ \hline
	Service 		& $10$ 				\\ \hline
	Theatre 		& $52$ 				\\ \hline
	Writing 		& $29$ 				\\ \hline
	\hline
	Total			& $461$				\\ \hline
	\end{tabular}
	\caption{Selected occupations obtained from the Wikipedia article \url{en.wikipedia.org/wiki/Lists_of_occupations}, grouped by category. We have selected a total of $453$ occupations from $11$ distinct groups (Artistic, Science, Service, etc.).}
\end{table}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{fontspec}
\usepackage{xeCJK}

\usepackage{multirow}
\usepackage{hyperref}
\usepackage{float}
\usepackage{array}
\usepackage{dirtytalk}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{graphicx}
%
%
%\usepackage{xltxtra}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|m{3cm}|c|}
	\hline
	Language Family & Language & Pronominal Gender System & Supported \\ \hline \hline
	\multirow{2}{*}{Austronesian} & Malay & No & Yes \\
	& Tagalog & No & \textbf{No} \\ \hline
	\multirow{3}{*}{Finno-Ugric} & Estonian & No & Yes \\
	& Finnish & No & Yes\\
	& Hungarian & No & Yes \\ \hline
	\multirow{4}{*}{Indo-European} & Armenian & No & Yes \\
	& Bengali & No & Yes \\
	& English & \textbf{Yes} & Yes \\
	& Persian & \textbf{Yes} & Yes \\ \hline
	\multirow{3}{*}{Indo-Aryan} & Maithili & No & \textbf{No} \\
	& Nepali & No & Yes \\
	& Oriya & No & \textbf{No} \\ \hline
	\multirow{10}{*}{} & Japanese & No & Yes \\
	& Korean & \textbf{Optional} & Yes \\
	& Turkish & No & Yes \\
	& Yoruba & No & Yes \\
	& Basque & No & Yes \\
	& Swahili & No & Yes \\
	& Chinese & \textbf{Optional} & Yes \\
	& Cantonese & No & \textbf{No} \\
	& Pipil & No & \textbf{No} \\
	& Quechuan & No & \textbf{No} \\ \hline
	\multirow{4}{*}{Constructed} & Esperanto & \textbf{Yes} & Yes \\
	& Ido & \textbf{Optional} & \textbf{No} \\
	& Lingua Franca Nova & No & \textbf{No} \\
	& Interlingua & No & \textbf{No} \\ \hline
	\end{tabular}
	\caption{Selected gender neutral languages obtained from the Wikipedia article \url{en.wikipedia.org/wiki/Gender_neutrality_in_genderless_languages}. Languages are grouped according to language families and classified according to whether they exhibit pronominal gender system. For the purposes of this work, we have decided to work only with languages lacking such a system, and as such Persian and Esperanto have been discarded. Because Tagalog, Maithili, Oriya, Cantonese, Pipil, Quechuan, Lingua Franca Nova and Interlingua lack support from Google Translate, these languages have also been ommitted from this work.}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
	\hline
	Category 		& \# Occupations 	\\ \hline \hline
	Artistic 		& $102$ 			\\ \hline
	Computer 		& $19$ 				\\ \hline
	Corporate 		& $50$ 				\\ \hline
	Dance 			& $9$ 				\\ \hline
	Film/Television & $26$ 				\\ \hline
	Healthcare 		& $88$ 				\\ \hline
	Industrial 		& $26$ 				\\ \hline
	Science 		& $50$ 				\\ \hline
	Service 		& $10$ 				\\ \hline
	Theatre 		& $52$ 				\\ \hline
	Writing 		& $29$ 				\\ \hline
	\hline
	Total			& $461$				\\ \hline
	\end{tabular}
	\caption{Selected occupations obtained from the Wikipedia article \url{en.wikipedia.org/wiki/Lists_of_occupations}, grouped by category. We have selected a total of $453$ occupations from $11$ distinct groups (Artistic, Science, Service, etc.).}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
	\hline
	Language 	& Sentence template \\ \hline \hline
	Malay		& dia adalah $\langle occupation \rangle$ \\ \hline
	Estonian	& ta on $\langle occupation \rangle$ \\ \hline
	Finnish		& hän on $\langle occupation \rangle$ \\ \hline
	Hungarian	& ő $\langle occupation \rangle$ \\ \hline
	Armenian	& նա $\langle occupation \rangle$ է \\ \hline
	Japanese	& は $\langle occupation \rangle$ です\\ \hline
	Turkish		& o bir $\langle occupation \rangle$ \\ \hline
	Yoruba		& o jẹ $\langle occupation \rangle$ \\ \hline
	Basque		& $\langle occupation \rangle$ da \\ \hline
	Swahili		& yeye ni $\langle occupation \rangle$ \\ \hline
	Chinese		& ta $\langle occupation \rangle$ \\ \hline
	\end{tabular}
	\caption{Templates used to infer gender biases in the translation to the English language.}
\end{table}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	Category & Female & Male & Neutral & Ratio & Total \\ \hline
	\hline
	Artistic        & 179    & 504  & 206     & 2.816 	& 918 \\ \hline
	Computer        & 7      & 125  & 39      & 17.857 	& 171 \\ \hline
	Corporate       & 36     & 340  & 74      & 9.444 	& 450 \\ \hline
	Dance           & 31     & 33   & 17      & 1.064 	& 81  \\ \hline
	Film-television & 54     & 125  & 55      & 2.315 	& 234 \\ \hline
	Healthcare      & 174    & 475  & 130     & 2.730 	& 792 \\ \hline
	Science         & 34     & 357  & 59      & 10.500 	& 450 \\ \hline
	Service         & 5      & 63   & 22      & 12.600 	& 90  \\ \hline
	Theatre         & 75     & 296  & 106     & 3.947 	& 477 \\ \hline
	Writing         & 41     & 148  & 72      & 3.610 	& 261 \\ \hline
	\end{tabular}
	\label{tab:gender-by-category}
	\caption{Number of female, male and neutral pronominal genders per occupation category in the translated sentences. The corresponding sex ratios (\# Male / \# Female) show just how much male defaults are prominent in male dominated fields such as computer science, with up to $\approx 18$ occurrences of male pronouns for each of a female one.}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pictures/gender-by-category}
	\label{fig:gender-by-category}
	\caption{The distribution of pronominal genders in the translated sentences not only suggests a tendency towards male defaults but also reflects the structure of male dominated fields, with the proportion of male pronouns amounting to $\approx 73\%$ in computer related jobs and $\approx 76\%$ in corporate jobs respectively. Because Google Translate occasionally fails to translate a sentence, the bars for some categories fail to add up to $100\%$.}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pictures/gender-by-language}
	\label{fig:gender-by-language}
	\caption{The distribution of pronominal genders in the translated sentences not only suggests a tendency towards male defaults but also reflects the structure of male dominated fields, with the proportion of male pronouns amounting to $\approx 73\%$ in computer related jobs and $\approx 76\%$ in corporate jobs respectively. Because Google Translate occasionally fails to translate a sentence, the bars for some categories fail to add up to $100\%$.}
\end{figure}
